MS_data_clean

### 2021-10-15 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep.Rmd
Refilters fowler so its not filtered only to GLOBI bees. Adds the pollen hosts as listed by GLOBI (to be compared to pollen hosts in MS data). Reruns tables to reflect changes.

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)

@@

```{r, echo = False}
beeCiteW
```

@@

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(freq)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = freq, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(freq))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(freq)) ## min of 5 citations to contribute to degree

@@

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee", type="inner") %>% filter(n > 5)

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee", type="inner") %>% filter(n > 5)
beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee", type="inner") %>% filter(n > 5)
```

@@

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0
```

@@

write.csv(plotdf, 'BeeCiteRaw')
```
```{r}
plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

>>>>

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/interactions_dup.csv') %>% select(-X)

@@

@@

TotalCitations <- beeCiteNoFilter %>% group_by(interestBee) %>% summarise(n_Bee = sum(n)) %>% ungroup()
TotalCitations <- sum(TotalCitations$n_Bee)

beeCiteNoFilter <- beeCiteNoFilter %>% group_by(interestBee) %>% mutate(interact= interact, n = n, n_Bee = sum(n)) %>% ungroup()
beeCiteNoFilter <- beeCiteNoFilter %>% summarise(interestBee = interestBee, interact= interact, n = n, n_Bee = n_Bee, BeeFreq = n_Bee / TotalCitations) %>% ungroup()

beeIntDegree <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegree <- beeIntDegree %>% group_by(interestBee) %>% summarise(degree = sum(n))

beeIntDegreeC10 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC10$numcitations <- beeCiteNoFilter$n
beeIntDegreeC10 <- beeIntDegreeC10 %>% filter(numcitations >= 10)
beeIntDegreeC10 <- beeIntDegreeC10 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 10 citations to contribute to degree

beeIntDegreeC5 <- beeCiteNoFilter %>% select(interestBee, interact) %>% unique() %>% group_by(interestBee, interact) %>% count()
beeIntDegreeC5$numcitations <- beeCiteNoFilter$n
beeIntDegreeC5 <- beeIntDegreeC5 %>% filter(numcitations >= 5)
beeIntDegreeC5 <- beeIntDegreeC5 %>% group_by(interestBee) %>% summarise(degree = sum(n)) ## min of 5 citations to contribute to degree

@@

beeCite <-  merge(beeCiteNoFilter, beeIntDegree, by= "interestBee", type = 'inner')

beeCiteC10 <-  merge(beeCiteNoFilter, beeIntDegreeC10, by= "interestBee",  type = 'inner')

beeCiteC5 <-  merge(beeCiteNoFilter, beeIntDegreeC5, by= "interestBee",  type = 'inner')

beeCiteW <-  merge(beeCiteNoFilter, beeIntDegreeW, by= "interestBee",  type = 'inner')

beeCiteW
```

@@

tablejoining <- merge(x = western_test, y = eastern_test, by = "interestBee", all = TRUE)
table_test <- merge(x = tablejoining, y = central_test, by = "interestBee", all = TRUE)
table_test[is.na(table_test)] <- 0

table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
table_test$specialist <- as.integer(table_test$specialist > 0)
table_test$specialist[is.na(table_test$specialist)] <- 0

write.csv(table_test, 'fowler')
```

@@

write.csv(plotdf, 'BeeCiteRaw')
```

```{r}
plotdf <- merge(beeCiteC5, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

### 2021-10-15 MS_data_modeling/MS_data_explore.Rmd
Refilters fowler so its not filtered only to GLOBI bees. Adds the pollen hosts as listed by GLOBI (to be compared to pollen hosts in MS data). Reruns tables to reflect changes.

fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/modeling_data/BeeCiteRaw') %>%
   mutate(interestBee = str_to_title(interestBee)) %>%
  select(interestBee, west_specialist, east_specialist, cent_specialist, specialist)
```

@@

```{r}
globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
  select(interestBee, degreeGLOBEE) %>% unique() %>%
  arrange(interestBee)

@@

compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS*existsFOWLER)
```

@@

compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE)
```

@@

## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa)

## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables)

>>>>

fowler <- read.csv('/Users/nickbachelder/Desktop/Bee-Specialization-Modeling/Fowler-GLOBI Modeling/fowler') %>%
   mutate(interestBee = str_to_title(interestBee)) %>%
  select(interestBee, west_specialist, east_specialist, cent_specialist, specialist)
```

@@

```{r}
pollenHostGLOBEE <- aggregate(interact~interestBee, globee_5cut, FUN=toString) %>% mutate(interestBee = str_to_title(interestBee)) %>% 
  rename('pollenHostGLOBEE' = interact)
## this is plant interactions for each bee genus according to globi

globee_format <- globee %>% mutate(interestBee = str_to_title(interestBee)) %>% rename(degreeGLOBEE = degree) %>% 
  select(interestBee, degreeGLOBEE) %>% unique() %>%
  arrange(interestBee) %>% merge(pollenHostGLOBEE, by = 'interestBee')

@@

compare_df <- compare_df %>% merge(fowler %>% rename(specialistFOWLER = specialist) %>% 
                                     select(interestBee, specialistFOWLER), by = c('interestBee'), all = T) %>% unique() %>%
  mutate(existsGLOBEE = ifelse(is.na(degreeGLOBEE), 0, 1), existsMS = ifelse(is.na(degreeMS), 0, 1), 
         existsFOWLER = ifelse(is.na(specialistFOWLER), 0, 1),
         existsALL = existsGLOBEE*existsMS*existsFOWLER,
         specialistFOWLER = ifelse(is.na(specialistFOWLER), 0 , specialistFOWLER),
         pollenHostGLOBEE = str_to_title(pollenHostGLOBEE)) %>% rename(pollenHostsMS = pollenHosts,
                                                                                              ReferencesMS = References) %>%
  select(interestBee, specializationMS, degreeGLOBEE, degreeGLOBEE_C5, degreeGLOBEE_T, specialistFOWLER, 
         existsGLOBEE, existsMS, existsFOWLER, existsALL, pollenHostsMS, pollenHostGLOBEE, ReferencesMS)

compare_df %>% filter(existsALL == 1)
```

@@

compare_df %>% filter(specialistFOWLER == 1, specializationMS  == 'Polylectic') %>% 
  select(interestBee, specializationMS, specialistFOWLER, degreeGLOBEE_C5, degreeGLOBEE, pollenHosts) %>% filter(degreeGLOBEE_C5 > 3)
```

@@

## Create slides / tables explaining odd bees between datasets (ie. fowler specialist vs MS non-specialist and vis versa) ## WIP

## Check how many GLOBI bees were cut when filtering for fowler, add them back (update previously done tables). ## DONE: Turns out that we didnt cut globi bees, but we instead cut FOWLER bees.

### Also: Added the target pollen hosts from globi so we can compare to pollen hosts from MS

```{r}
compare_df
```
